---
title: "Machine Learning Project"
author: "Marius Nicolescu"
date: "19 octombrie 2015"
output: html_document
---

Executive summary
-----------------


Using more types devices it is now possible to collect a large amount of data about personal activity. Data was collected from 6 participants, from accelerometers mounted on the arm, belly on the belt, forearm, arm, and dumbbell. 

Our goal is to develop a machine learning to predict the type of movement that do a participant. For this we will use three model.  

The first and second will be based RandomForest (but using different packages of R caret and RandomForest) and the third model predict wiht tree. 

Finally we compare the performance of the three models (the speed and accuracy)


Load data
----------

The data for this assignment comes from (http://groupware.les.inf.puc-rio.br/har) and contains information from accelerometers.

The data are split into a training group  this  have  19.622 rows, 160 columns and test set( for predict) 20 rows  and 160 columns.

Data from these sensors are associated with the following five activities: Sitting, Sitting down, Standing, Standing up, Walking. The outcome used for this is the ???classe??? variable.

```{r, results="hide", warning=FALSE}

library(caret)
library(randomForest)
library(rpart)
library(rpart.plot)
##Load data from file
setwd("~/R/predict")
datatrain0<-read.csv("pml-training (2).csv", na.strings = c(NA,"","DIV/0!"))
datatest<-read.csv("pml-testing (1).csv")
```

Remove the columns not needed.
------------------------------

Some columns contain Nas (empy data), will remove columns that contain Nas. Also we remove the first 7 columns that contain information that is not  useful for prediction. The last column contains outliner, the type of movement that makes the person.
```{r}
## Cleaning data 
datatrain1<-datatrain0[,colSums(is.na(datatrain0))==0]
datatest1<-datatest[,colSums(is.na(datatest))==0]
datatrain1<-datatrain1[,-(1:6)]
```

Now  dimension of traing set is:   
```{r, echo=FALSE}
dim(datatrain1)
```

Create the cross validation trainset
----------------------------------
The training data is split 70:30 for cross validation. 70% is used for training and 30% is used for validation.
```{r}
#Preproces trainig test
intrain<-createDataPartition(y=datatrain1$classe,p=0.7,list=FALSE)
training2<-datatrain1[intrain,]
testing2<-datatrain1[-intrain,]
```


Train Model
-----------

I tried to use that training and regression model multiiniara, but the results were weak (acurracy = 0.34). RandomForest method is more recommended.

```{r}
# Matrix for record performance
performance<-matrix(data=NA,nrow = 2,ncol=3)
colnames(performance)<-c("RandomForest(RandomForest)","RandomForest(Caret)","Tree(rpart)")
row.names(performance)<-c("Time/Speed", "Accuracy")

## Model1 Random Forest with package randomForest
procesortime<-proc.time()
model1<-randomForest(classe~.,data=training2,methods="class")
t<-proc.time()-procesortime
performance[1,1]<-t[1]
#summary(model1)
prediction1<-predict(model1,testing2,type="class")
Matrix1<-confusionMatrix(prediction1,testing2$classe)
performance[2,1]<-Matrix1$overall[1]
# Model 2 Random Forest with package caret 
procesortime<-proc.time()
model2<- train(classe~., method="rf", data=training2, 
                 trControl=trainControl(method='cv',5), ntree=250)
t<-proc.time()-procesortime
performance[1,2]<-t[1]
predict2<-predict(model2, testing2)
Matrix2<-confusionMatrix(predict2,testing2$classe)
performance[2,2]<-Matrix2$overall[1]
 #Model 3 Tree  package rpart
procesortime<-proc.time()
treeView<-rpart(classe~., data=training2 , method="class")
t<-proc.time()-procesortime
performance[1,3]<-t[1]
predict3<-predict(treeView, testing2,type="class")
Matrix3<-confusionMatrix(predict3,testing2$classe)
performance[2,3]<-Matrix3$overall[1]
```


First model 
----------

With many variables dependent on each other, Random Forest is a recommended model
 Fist model use Random Forest packet 
Type of random forest: classification, Number of trees: 500, each split: 7

```{r, echo=FALSE}
Matrix1
```

Second Model
------------

Random Forest packet Caret

```{r, echo=FALSE}
Matrix2

```

Third Model
------------
Predict with tree 

```{r,echo=FALSE}
Matrix3

```

Performance
------------

Performance for three models are:
```{r,echo=FALSE}
performance

```

View Tree 

```{r,ech=FALSE, fig.height=7}
prp(treeView)

```


Conclusion and Final Test
-------------------------

As we assumed, RandomForest method has the best accuracy, if we use the method is faster RandomForest package it is the fastest.

The model shows a high prediction accuracy of 99.8 % in the cross-validation.

The model obtained a 100% prediction accuracy of the 20 cases in the test set.
